### 2016-02-25 ch02-qlogms/excitation-design.tex
Some minor changes to CH02

\item Robust input design one tries to make the input design robust to uncertainty in the prior knowledge of the parameter values, see e.g. \citep{Rojas2007} and references therein. 
While the signal is more robust towards uncertainty in the prior knowledge, it requires more forethought than (nominal) optimal designs.
E.g. the approach followed by \citet{Rojas2007} relies on both nominal parameter values and  knowledge of the distribution function of these parameters to design a robust (optimal) excitation signal.
As such, this requires more advanced prior knowledge of the system to cope with the uncertainty of the prior knowledge.
\end{enumerate}

@@

  \paragraph{Contents}
  In \secref{sec:excitation:inputDesign}, we first consider (optimal) input design for an example second-order system.
  Next in \secref{}, \gls{LTI} systems are parametrized to facilitate the input design.
  This will lead to signals with a $1/f$ power spectrum as in \secref{sec:excitation:logarithmic}.
  In \secref{sec:excitation:multisine} a comparison of different discrete frequency grids is provided, this is afterwards related to the problem of generating a multisine signal with a $1/f$ power spectral density.
  Next, \secref{sec:excitation:optimAlpha} describes the selection of a suitable frequency ratio $\alpha$ for a logarithmically spaced grid such that constant model quality is obtained.
  The usefulness of the proposed signals is illustrated on simulations in \secref{sec:excitation:simulation} and validated on the measurement of a set of band-pass filters in \secref{sec:excitation:measurement}.
  Conclusions are formulated in \secref{sec:excitation:conclusion}.

@@

Alternatively, one could also construct a signal with power spectrum concentrated at discrete frequency lines.
Such signals are elaborated in \secref{sec:excitation:method}.

@@

Moreover, the results from \secref{sec:excitation:ex:sumOfSecondOrderSipmle} for the sum of simplified second order systems, i.e. equation \eqref{eq:excitation:sumSys}, extend directly to the generic \gls{LTI} system in \eqref{eq:excitation:PFE} for the second-order sub-systems.

@@

\subsection{Equidistant (Linear) Grid Multisine}
  In many applications, an equidistant frequency grid is used for $u\left( t\right) $ because this fits best with the classical DFT analysis \citep{OppenheimDT,Mandal2007}. 
  It allows to analyze periodic signals on an equidistant frequency grid very efficiently.
  For such a frequency grid, it is guaranteed that the signal is periodic in the time domain, unlike for the generalized multisines of the previous section.

@@

  Using such linear grid, we can easily construct a representation of a multisine
  in discrete time that is suited for use with the DFT by sampling at a given sampling rate $f_s$:
  \begin{equation}
     u \left[ n\right] = \frac{1}{\sqrt{F}}
     \sum_{k=1}^{F} 
       A_k 
       \sin 
         \left(\frac{2\pi n \left( k + k_0 \right) \Delta f}{f_s} + \phi_k \right)
    \label{eq:excitation:MultiSineDT}
  \end{equation}

@@

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}
  \begin{example}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}

@@

  For that reason, a logarithmic multisine is a good candidate for our purposes.
  However, it is more involved both to generate and to analyze such a signal, as the excited frequencies do not lie on a commensurate frequency grid.
  Especially for periodic measurements, this is a major downside since the period of such signal does not always exist (i.e. this leads to infinite measurement times).
  It is therefore not practical to use a perfectly logarithmically spaced frequency grid as the \gls{DFT} or \gls{CZT}~\citep{Rabiner2004} explicitly rely on a commensurate frequency grid.

@@

  This allows to define the amplitude spectrum $A_k$ as
  \begin{equation}
    A_k = A_{\mathrm{in}}
                 \sqrt{\int_{\fMin{k}}
                     ^{\fMax{k}}
                     \limits
                     \frac{1}
                          {f}
                     \,\dd{f}}
          = A_{\mathrm{in}}
                \cdot
                \sqrt{
                \ln \frac{\;\fMax{k}\;}{\fMin{k}}
                }
  \end{equation}
  where $\left[ \fMin{k}, \fMax{k} \right] $ denote the frequency range over which PSD of the pink noise is to be approximated by the power at frequency line $f_k$.

@@

  To determine a suitable value of the frequency spacing ratio $\alpha$ for the logarithmic and (compensated) quasi-logarithmic grid, we construct a logarithmic multisine signal in the frequency domain for a given frequency band (in this example, $\left[ 0.01 \unit{rad/s}, 100 \unit{rad/s}\right]$ was used).
  In the previous section we saw that systems with equal damping and DC gain, will behave very similarly when excited by a log signal with spacing $\alpha$.

@@

  We choose the resonance frequency of the systems such that
  \begin{equation}
   \wM\left( \wn \right)  = \omega_k + \beta \left( \omega_{k+1} - \omega_k \right)
   \text{ with } 0 \leq \beta < 1
   \label{eq:excitation:chosenVariationOmega}
  \end{equation}
  where $\omega_k$ and $\omega_{k+1}$ are the excited grid lines closest to the frequency $\wM$ where the FRF has its peak amplitude and $\beta$ is the misalignment coefficient that is varied over the interval $\RightOpenInterval{0,1}$ as depicted graphically in \figref{fig:excitation:misalignment-logms}
  This means that $\omega_k \leq \wM \leq \omega_{k+1}$ and that $\beta$ replaces the $\gamma$ factor in \eqref{eq:excitation:theoreticalBestVariationOmega}.
  If we take a look at the special case when $\beta=0$ (or $\beta=1$) in this expression, we see that this situation occurs when the peak value of the FRF at $\wM$ coincides with the grid line $\omega_k$ or $\omega_{k+1}$ respectively.
  Intuitively, we study second order systems whose peak frequency $\wM$ is shifted between two neighboring excited frequency lines.
  From the previous section, we know that the exact frequency matters little as long as it remains within the bulk of the frequency grid.
  In this particular example $\wn$ was chosen at approximately $1 \unit{rad/s}$.

  We normalize the root mean square (RMS) value of the excitation signal to $1$ and choose $\sigma_E\left( \omega \right) = 0.1$ for all evaluations.
  The noise level is kept constant, which causes the \gls{SNR} per excited line to drop for more finely spaced excitation signals.
  % SNR correct: Pnoise ~ sE^2
  %              Psignal ~F*U^2 = 1 ==> U^2 = 1/F
  %              SNR = Psignal / Pnoise = 1/(sE^2 * F)
  %          ==> SNR decreases as F increases

  We then calculate the theoretical variance on the magnitude of the peak in the FRF ($\left| G\left( \wM \right) \right|$) as shown in \eqref{eq:excitation:varPeak} for a coarse grid.
  A coarse grid was chosen with $\tilde{\alpha} = 10$; this means that the spacing between the excited lines around the resonance are much larger than the system bandwidth.
  It is obvious that such an excitation signal will not perform well in most cases, as the resonance peak is very likely to be overlooked in such an \gls{FRF}.

@@

  \begin{figure}
    \centering
      \setlength\figurewidth{0.68\columnwidth}
      \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/betaWC.tikz}
    \caption[Variance $\sigma_G^2\left( \wM \right)$ as a function of the grid misalignment coefficient $\beta$.]{Evaluation of the variance $\sigma_G^2\left( \wM \right) $ for different prototype systems as function of the misalignment $\beta$.
             The full lines depict the variance for a coarsely spaced signal.
             The dashed lines show the variance for a signal with the optimal $\alpha = \alpha^{\star}$ (cfr. \eqref{eq:excitation:optimAlpha}).}
    \label{fig:excitation:worstCaseBeta}
  \end{figure}

@@

  For a large peak amplitude of the FRF (low damping), the worst case increase (marked with a cross) is approximately centered around $\beta = 0.5$ (i.e. midway between the nearest excited lines).
  In systems with a higher damping, this symmetry is lost.
  This can be explained intuitively as the $3\unit{dB}$ bandwidth is less symmetric for highly damped systems than in lightly damped systems with a very narrow resonance peak.
  Also note that the value of $\sigma_G\left(\wM \right)$ is practically identical for $\beta=0$ and $\beta=1$.
  This is in accordance with the claim that in the bulk of the grid, a shift over an integer number of lines will only marginally influence the uncertainty.

@@

  To give a more intuitive feeling how the frequency spacing $\alpha$ and the system under test react, both the input spectrum and the FRF of the system are visualized in \figref{fig:excitation:signalAndSystem} for the worst value of $\beta$ and the optimal frequency spacing $\alpha^{\star}$.
  The $3\unit{dB}$ bandwidth of the system is highlighted in gray together with the excited frequency lines that lie within that bandwidth.

@@

\section{Simulations} \label{sec:excitation:simulation}
In the following set of simulations, we will study how the FRF of different second order systems with equal damping (and the corresponding uncertainty) are measured by the multisine excitation signals from \secref{sec:excitation:method}.
  \subsection{Setup}
  We consider second-order systems with an equal damping and with resonance frequencies $4$, $8$, $20$, $40$, $80$, $200$, $400$, $800$, $2000$ and $4000 \unit{Hz}$.
  Using the \matlab\ instruction \texttt{cheby1}, discrete-time equivalents of these systems are generated.
  These can be represented by a transfer function of the form:
  \begin{equation}
    G \left( z \right)\Big|_{\damping, \fR} =
       \frac{         b_2  z^{2} + b_1 z + b_0}%
            {\phantom{a_2} z^{2} + a_1 z + a_0}
  \text{.}
  \label{eq:excitation:secondOrderSystemDTidentifiable}
  \end{equation}

  This form can also be obtained by applying Tustin's method or bilinear transform \citep{Oppenheim1983} to the continuous time system \eqref{eq:excitation:secondOrderSystemCT}.

  The input signal has an equidistant frequency grid with a resolution $\Delta f = 1\unit{Hz}$ between $1 \unit{Hz}$ and $16\,384\unit{Hz}$.
  The sampling frequency $f_s$ is set to $65\,536\unit{Hz}$ to prevent aliasing.
  For the quasi-log grid multisine, the compensated and pink quasi-log multisine, we chose a frequency ratio $\alpha = 1.05$.
  According to \eqref{eq:excitation:optimAlpha}, this frequency ratio is suited for second-order systems with a damping $\damping \geq 0.05$, or a peak amplitude of the FRF $\abs{G(\wM)}$ less than approximately $20\unit{dB}$.
  Each of these signals is constructed to have an RMS value of $1$.
  The amplitude spectra of these signals are displayed in \figref{fig:excitation:freqGrids} for identical signal power.

@@

  The obtained FRF is then fitted to the transfer function model \eqref{eq:excitation:secondOrderSystemDTidentifiable} using the \texttt{elis} function of the \FDIDENT~\citep{FDIDENT} for \MATLAB.
  This is a maximum likelihood estimator that fits a parametric model to the measured data in the frequency domain.
  % \TODO{add formulas with cost function etc.}
  To aid convergence and avoid local minima, the initial values for the model are chosen equal to the parameters of the true system $G_0(z)$, so that we can focus completely on the variance of the estimations.

@@

  \begin{figure}[th]
    \centering
      \setlength\figurewidth{0.68\columnwidth}
      \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/Sys02-4.tikz}
    \caption[Simulated FRF and its variance for different excitation signals.]{\gls{FRF}~\legref{leg:excitation:sim:FRF} of a second order system with $\fR=40\unit{Hz}$ and $\damping=0.2$ along with the corresponding variance and bias of the parametric models obtained using three different kinds of multisine excitations.}%
    \label{fig:excitation:FRF1}
  \end{figure}

@@

  \begin{figure}%
    \centering
      \setlength\figurewidth{0.68\columnwidth}
      \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/damping02-cmp.tikz}
    \caption[Simulated FRFs and their variances of systems with $\damping=0.2$ for different excitation signals.]{\glspl{FRF} for different second order systems with constant damping
             $\damping=0.2$ but different resonance frequencies. Comparison of the
             variances around the respective resonance frequency for the different
             excitation signals.}%
    \label{fig:excitation:damping02}
  \end{figure}

@@

  \subsection{Systems with a lower damping}
  Given the optimal frequency ratio \eqref{eq:excitation:optimAlpha}, one expects that quasi-logarithmic grids with a frequency ratio $\alpha = 1.05$ will allow to identify second order systems with damping $\damping = 0.05$ with a constant uncertainty.
  In this set of simulations, we check this by repeating the previous set for such systems.
  In this case, the output noise is chosen $\sigma_e \approx 0.1123$ to keep the SNR of $20\unit{dB}$.
  The corresponding results are shown in \figref{fig:excitation:damping005}.

  \begin{figure}%
    \centering
      \setlength\figurewidth{0.68\columnwidth}
      \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/damping005-cmp.tikz}
    \caption[Simulated FRFs and their variances of systems with $\damping=0.05$ for different excitation signals.]{FRFs for different second order systems with constant damping
             $\damping=0.05$ but different resonance frequencies. Comparison of the
             variances around the respective resonance frequency for the different
             excitation signals.}%
    \label{fig:excitation:damping005}
  \end{figure}

@@

    $R=50$ periods of the multisine were measured, from which the first period was discarded to suppress transient effects.
    For each of the remaining periods, the FRF of the filter is calculated and its average $G\left(\omega \right)$ over the periods is calculated.
    We also calculate the corresponding standard deviation $\sigma_G\left( \omega \right)$ of the FRF and normalize it for a single period of the excitation signal.
    We denote $G\left(\omega \right)$ the non-parametric FRF and $\sigma_G\left( \omega \right)$ its standard deviation, as these measures are determined completely nonparametrically from the measurements.

@@

    \subsection{Results}
    In \figref{fig:excitation:measurements} the nonparametric FRF and the standard deviation of the corresponding parametric estimates are shown in a similar way as for the simulations in the previous section.

@@

We have also shown an effective way to choose the frequency spacing of a (quasi)-logarithmically spaced multisine for an equidamped set of systems.
The presented method allows to select a maximal value for the frequency ratio $\alpha = 1 + \damping$ for which the performance will not suffer from the unavoidable misalignment of the system with respect to the excited frequencies.
In practice we advise to use an excitation signal with approximately four excited lines within the $3\unit{dB}$ bandwidth of each second-order subsystem.

>>>>

\item Robust input design one tries to make the input design robust to uncertainty in the prior knowledge of the parameter values, see e.g.~\citep{Rojas2007} and references therein. 
While the signal is more robust towards uncertainty in the prior knowledge, it requires more forethought than (nominal) optimal designs.
E.g. the approach followed by~\citet{Rojas2007} relies on both nominal parameter values and  knowledge of the distribution function of these parameters to design a robust (optimal) excitation signal.
As such, this requires more advanced prior knowledge of the system to cope with the uncertainty of the prior knowledge.
\end{enumerate}

@@

  \paragraph{Contents}
  In \secref{sec:excitation:inputDesign}, we first consider (optimal) input design for an example second-order system.
  \secref{sec:excitation:logarithmic} expounds the use of a signal with $1/f$  power spectrum on some basic sub-systems.
  Next, in \secref{sec:excitation:parametrization} \gls{LTI} systems are parametrized to facilitate a robust input design based on the observations of the preceding section.
  In \secref{sec:excitation:multisine} a comparison of different discrete frequency grids is provided, this is related to the problem of generating a multisine signal with a $1/f$ power spectral density.
  Next, \secref{sec:excitation:optimAlpha} describes the selection of a suitable frequency ratio $\alpha$ for a logarithmically spaced grid such that constant model quality is obtained.
  The usefulness of the proposed signals is illustrated on simulations in \secref{sec:excitation:simulation} and validated on the measurement of a set of band-pass filters in \secref{sec:excitation:measurement}.
  Conclusions are formulated in \secref{sec:excitation:conclusion}.

@@

Alternatively, one could also construct a signal with power spectrum concentrated at discrete frequency lines.
Such signals are elaborated in \secref{sec:excitation:multisine}.

@@

Moreover, the results from \secref{sec:excitation:ex:sumOfSecondOrderSipmle} for the sum of simplified second order systems, i.e. equation \eqref{eq:excitation:sumSys}, extend directly to the generic \gls{LTI} system in \eqref{eq:excitation:PFE} for the second-order sub-systems.
\begin{assumption}

\end{assumption}

@@

\subsection{Equidistant (Linear) Grid Multisine}
  In many applications, an equidistant frequency grid is used for $u\left( t\right) $ because this fits best with the classical \gls{DFT} analysis \citep{OppenheimDT,Mandal2007}. 
  It allows to analyze periodic signals on an equidistant frequency grid very efficiently.
  For such a frequency grid, it is guaranteed that the signal is periodic in the time domain, unlike for the generalized multisines of the previous section.

@@

  Using such linear grid, we can easily construct a representation of a multisine
  in discrete time that is suited for use with the \gls{DFT} by sampling at a given sampling rate $f_s$:
  \begin{equation}
     u \left[ n\right] = \frac{1}{\sqrt{F}}
     \sum_{k=1}^{F} 
       A_k 
       \sin 
         \left(\frac{2\pi n \left( k + k_0 \right) \Delta f}{f_s} + \phi_k \right)
    \label{eq:excitation:MultiSineDT}
  \end{equation}

@@

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

@@

  For that reason, a logarithmic multisine is a good candidate for our purposes.
  However, it is more involved both to generate and to analyze such a signal, as the excited frequencies do not lie on a commensurate frequency grid.
  Especially for periodic measurements, this is a major downside since the period of such signal does not always exist (i.e. this leads to infinite measurement times).
  It is therefore not practical to use a perfectly logarithmically spaced frequency grid as the \gls{DFT} or \gls{CZT}~\citep{Rabiner2004} explicitly rely on an equidistantly-spaced frequency grid.

@@

  This allows to define the amplitude spectrum $A_k$ as
  \begin{equation}
    A_k = A_{\mathrm{in}}
                 \sqrt{\int_{\fMin{k}}
                     ^{\fMax{k}}
                     \limits
                     \frac{1}
                          {f}
                     \,\dd{f}}
          = A_{\mathrm{in}}
                \cdot
                \sqrt{
                \ln \frac{\;\fMax{k}\;}{\fMin{k}}
                }
  \end{equation}
  where $\left[ \fMin{k}, \fMax{k} \right] $ denote the frequency range over which \gls{PSD} of the pink noise is to be approximated by the power at frequency line $f_k$.

@@

  To determine a suitable value of the frequency spacing ratio $\alpha$ for the logarithmic and (compensated) quasi-logarithmic grid, we construct a logarithmic multisine signal in the frequency domain for a given frequency band (in this example, $\left[ 0.01 \unit{rad/s}, 100 \unit{rad/s}\right]$ was used).
  In the previous section we saw that systems with equal damping and \gls{DC} gain, will behave very similarly when excited by a log signal with spacing $\alpha$.

@@

  We choose the resonance frequency of the systems such that
  \begin{equation}
   \wM\left( \wn \right)  = \omega_k + \beta \left( \omega_{k+1} - \omega_k \right)
   \text{ with } 0 \leq \beta < 1
   \label{eq:excitation:chosenVariationOmega}
  \end{equation}
  where $\omega_k$ and $\omega_{k+1}$ are the excited grid lines closest to the frequency $\wM$ where the FRF has its peak amplitude and $\beta$ is the misalignment coefficient that is varied over the interval $\RightOpenInterval{0,1}$ as depicted graphically in \figref{fig:excitation:misalignment-logms}
  This means that $\omega_k \leq \wM \leq \omega_{k+1}$ and that $\beta$ replaces the $\gamma$ factor in \eqref{eq:excitation:theoreticalBestVariationOmega}.
  If we take a look at the special case when $\beta=0$ (or $\beta=1$) in this expression, we see that this situation occurs when the peak value of the \gls{FRF} at $\wM$ coincides with the grid line $\omega_k$ or $\omega_{k+1}$ respectively.
  Intuitively, we study second order systems whose peak frequency $\wM$ is shifted between two neighboring excited frequency lines.
  From the previous section, we know that the exact frequency matters little as long as it remains within the bulk of the frequency grid.
  In this particular example $\wn$ was chosen at approximately $1 \unit{rad/s}$.

  We normalize the\gls{RMS} value of the excitation signal to $1$ and choose $\sigma_E\left( \omega \right) = 0.1$ for all evaluations.
  The noise level is kept constant, which causes the \gls{SNR} per excited line to drop for more finely spaced excitation signals.
  % SNR correct: Pnoise ~ sE^2
  %              Psignal ~F*U^2 = 1 ==> U^2 = 1/F
  %              SNR = Psignal / Pnoise = 1/(sE^2 * F)
  %          ==> SNR decreases as F increases

  We then calculate the theoretical variance on the magnitude of the peak in the \gls{FRF} ($\left| G\left( \wM \right) \right|$) as shown in \eqref{eq:excitation:varPeak} for a coarse grid.
  A coarse grid was chosen with $\tilde{\alpha} = 10$; this means that the spacing between the excited lines around the resonance are much larger than the system bandwidth.
  It is obvious that such an excitation signal will not perform well in most cases, as the resonance peak is very likely to be overlooked in such an \gls{FRF}.

@@

  \begin{figure}
    \centering
      \setlength\figurewidth{0.68\columnwidth}
      \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/betaWC.tikz}
    \caption[Variance $\sigma_G^2\left( \wM \right)$ as a function of the grid misalignment coefficient $\beta$.]{Evaluation of the variance $\sigma_G^2\left( \wM \right) $ for different prototype systems as function of the misalignment $\beta$.
             The full lines depict the variance for a coarsely spaced signal.
             The dashed lines show the variance for a signal with the optimal $\alpha = \alpha^{\star}$ (see \eqref{eq:excitation:optimAlpha}).}
    \label{fig:excitation:worstCaseBeta}
  \end{figure}

@@

  For a large peak amplitude of the \gls{FRF} (low damping), the worst case increase (marked with a cross) is approximately centered around $\beta = 0.5$ (i.e. midway between the nearest excited lines).
  In systems with a higher damping, this symmetry is lost.
  This can be explained intuitively as the $3\unit{dB}$ bandwidth is less symmetric for highly damped systems than in lightly damped systems with a very narrow resonance peak.
  Also note that the value of $\sigma_G\left(\wM \right)$ is practically identical for $\beta=0$ and $\beta=1$.
  This is in accordance with the claim that in the bulk of the grid, a shift over an integer number of lines will only marginally influence the uncertainty.

@@

  To give a more intuitive feeling how the frequency spacing $\alpha$ and the system under test react, both the input spectrum and the \gls{FRF} of the system are visualized in \figref{fig:excitation:signalAndSystem} for the worst value of $\beta$ and the optimal frequency spacing $\alpha^{\star}$.
  The $3\unit{dB}$ bandwidth of the system is highlighted in gray together with the excited frequency lines that lie within that bandwidth.

@@

\section{Simulations} \label{sec:excitation:simulation}
In the following set of simulations, we will study how the \gls{FRF} of different second order systems with equal damping (and the corresponding uncertainty) are measured by the multisine excitation signals from \secref{sec:excitation:multisine}.
  \subsection{Setup}
  We consider second-order systems with an equal damping and with resonance frequencies $4$, $8$, $20$, $40$, $80$, $200$, $400$, $800$, $2000$ and $4000 \unit{Hz}$.
  Using the \matlab\ instruction \texttt{cheby1}, discrete-time equivalents of these systems are generated.
  These can be represented by a transfer function of the form:
  \begin{equation}
    G \left( z \right)\Big|_{\damping, \fR} =
       \frac{         b_2  z^{2} + b_1 z + b_0}%
            {\phantom{a_2} z^{2} + a_1 z + a_0}
  \text{.}
  \label{eq:excitation:secondOrderSystemDTidentifiable}
  \end{equation}

  This form can also be obtained by applying Tustin's method or bilinear transform~\citep{Oppenheim1983} to the continuous time system \eqref{eq:excitation:secondOrderSystemCT}.

  The input signal has an equidistant frequency grid with a resolution $\Delta f = 1\unit{Hz}$ between $1 \unit{Hz}$ and $16\,384\unit{Hz}$.
  The sampling frequency $f_s$ is set to $65\,536\unit{Hz}$ to prevent aliasing.
  For the quasi-log grid multisine, the compensated and pink quasi-log multisine, we chose a frequency ratio $\alpha = 1.05$.
  According to \eqref{eq:excitation:optimAlpha}, this frequency ratio is suited for second-order systems with a damping $\damping \geq 0.05$, or a peak amplitude of the \gls{FRF} $\abs{G(\wM)}$ less than approximately $20\unit{dB}$.
  Each of these signals is constructed to have an \gls{RMS} value of $1$.
  The amplitude spectra of these signals are displayed in \figref{fig:excitation:freqGrids} for identical signal power.

@@

  The obtained FRF is then fitted to the transfer function model \eqref{eq:excitation:secondOrderSystemDTidentifiable} using the \texttt{elis} function of the \FDIDENT~\citep{FDIDENT} for \MATLAB.
  This is a maximum likelihood estimator that fits a parametric model to the measured data in the frequency domain.
  %TODO: add formulas with cost function etc.
  To aid convergence and avoid local minima, the initial values for the model are chosen equal to the parameters of the true system $G_0(z)$, so that we can focus completely on the variance of the estimations.

@@

  \begin{figure}[th]
    \centering
      \setlength\figurewidth{0.68\columnwidth}
      \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/Sys02-4.tikz}
    \caption[Simulated \gls{FRF} and its variance for different excitation signals.]{\gls{FRF}~\legref{leg:excitation:sim:FRF} of a second order system with $\fR=40\unit{Hz}$ and $\damping=0.2$ along with the corresponding variance and bias of the parametric models obtained using three different kinds of multisine excitations.}%
    \label{fig:excitation:FRF1}
  \end{figure}

@@

  \begin{figure}%
    \centering
      \setlength\figurewidth{0.68\columnwidth}
      \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/damping02-cmp.tikz}
    \caption[Simulated \glspl{FRF} and their variances of systems with $\damping=0.2$ for different excitation signals.]{%
       \Glspl{FRF} for different second order systems with constant damping $\damping=0.2$ but different resonance frequencies. 
       Comparison of the variance around the respective resonance frequency reveals the performance of each excitation signal.}%
    \label{fig:excitation:damping02}
  \end{figure}

@@

  \subsection{Systems with a lower damping}
  Given the optimal frequency ratio \eqref{eq:excitation:optimAlpha}, one expects that quasi-logarithmic grids with a frequency ratio $\alpha = 1.05$ will allow to identify second order systems with damping $\damping = 0.05$ with a constant uncertainty.
  In this set of simulations, we check this by repeating the previous set for such systems.
  In this case, the output noise is chosen $\sigma_e \approx 0.1123$ to keep the \gls{SNR} of $20\unit{dB}$.
  The corresponding results are shown in \figref{fig:excitation:damping005}.

  \begin{figure}%
    \centering
      \setlength\figurewidth{0.68\columnwidth}
      \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/damping005-cmp.tikz}
    \caption[Simulated \glspl{FRF} and their variances of systems with $\damping=0.05$ for different excitation signals.]{\Glspl{FRF} for different second order systems with constant damping $\damping=0.05$ but different resonance frequencies. 
    Comparison of the variance around the respective resonance frequency reveals the performance of each excitation signal.}%
    \label{fig:excitation:damping005}
  \end{figure}

@@

    $R=50$ periods of the multisine were measured, from which the first period was discarded to suppress transient effects.
    For each of the remaining periods, the \gls{FRF} of the filter is calculated and its average $G\left(\omega \right)$ over the periods is calculated.
    We also calculate the corresponding standard deviation $\sigma_G\left( \omega \right)$ of the \gls{FRF} and normalize it for a single period of the excitation signal.
    We denote $G\left(\omega \right)$ the non-parametric FRF and $\sigma_G\left( \omega \right)$ its standard deviation, as these measures are determined completely nonparametrically from the measurements.

@@

    \subsection{Results}
    In \figref{fig:excitation:measurements} the nonparametric \gls{FRF} and the standard deviation of the corresponding parametric estimates are shown in a similar way as for the simulations in the previous section.

@@

We have also shown an effective way to choose the frequency spacing of a (quasi)-logarithmically spaced multisine for an equi-damped set of systems.
The presented method allows to select a maximal value for the frequency ratio $\alpha = 1 + \damping$ for which the performance will not suffer from the unavoidable misalignment of the system with respect to the excited frequencies.
In practice we advise to use an excitation signal with approximately four excited lines within the $3\unit{dB}$ bandwidth of each second-order subsystem.
